---
title: "Water_Quality_Assessment"
author: "Group H"
date: "2024-10-08"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
df = read.csv("/Users/vichu/Desktop/fda.csv",header=FALSE)
colnames(df) = df[1, ] 
df = df[-1, ] 
head(df,5)
```
```{r}
str(df)
```

```{r}
df$`Temperature Min` = as.numeric(df$`Temperature Min`)
df$`Temperature Max`= as.numeric(df$`Temperature Max`)
df$`pH Min` = as.numeric(df$`pH Min`)
df$`pH Max`= as.numeric(df$`pH Max`)
df$`Conductivity (µmhos/cm) Min`= as.numeric(df$`Conductivity (µmhos/cm) Min`)
df$`Conductivity (µmhos/cm) Max`= as.numeric(df$`Conductivity (µmhos/cm) Max`)
df$Year = as.numeric(df$Year)
df$`Station Name`=as.factor(df$`Station Name`)
df$STATE = as.factor(df$STATE)

df$`Station Code` = as.numeric(df$`Station Code`)

str(df)

```

```{r}
print(sum(is.na(df)))
```

```{r}
colSums(is.na(df))
```

```{r}
na_rows = df[!complete.cases(df),]
head(na_rows,5)
```


```{r}
library(zoo)
df$`Station Code` = na.approx(df$`Station Code`)
df$`Temperature Min` = na.approx(df$`Temperature Min`)
df$`Temperature Max`= na.approx(df$`Temperature Max`)
df$`pH Min` = na.approx(df$`pH Min`)
df$`pH Max` = na.approx(df$`pH Max`)
df$`Conductivity (µmhos/cm) Min` = na.approx(df$`Conductivity (µmhos/cm) Min`)
df$`Conductivity (µmhos/cm) Max` = na.approx(df$`Conductivity (µmhos/cm) Max`)
df$Year = na.approx(df$Year)
```
```{r}
colSums(is.na(df))
```
```{r}
str(df)
```

```{r}
library(dplyr)
library(stringr)
library(tm)        
library(SnowballC)
preprocess_text <- function(text) {

  text <- tolower(text)

  text <- str_replace_all(text, "[^\\w\\s]", "")

  text <- str_replace_all(text, "\\d+", "")

  text <- str_trim(text)
 
  tokens <- unlist(str_split(text, "\\s+"))

  stop_words <- stopwords("en")
  tokens <- tokens[!tokens %in% stop_words]

  tokens <- wordStem(tokens)

  return(paste(tokens, collapse = " "))
}


df$`Station Name` <- as.character(df$`Station Name`)
df$STATE <- as.character(df$STATE)

df <- df %>%
  mutate(
    processed_station_name = sapply(`Station Name`, preprocess_text),
    processed_state = sapply(STATE, preprocess_text)
  )

head(df,5)

```
```{r}
str(df)
```

```{r}
numeric_columns <- df %>%
  select(`Temperature Min`, `Temperature Max`, `pH Min`, `pH Max`, `Conductivity (µmhos/cm) Min`, `Conductivity (µmhos/cm) Max`)

correlation_matrix <- cor(numeric_columns, use = "complete.obs")
print(correlation_matrix)
```

```{r}
library(corrplot)

corrplot(correlation_matrix, method = "color", tl.col = "black", tl.srt = 45, addCoef.col = "black", number.cex = 0.7)

```

```{r}
library(dplyr)
yearly_data <- df %>%
  group_by(Year, `Station Name`, STATE) %>%
  summarise(
    Avg_Temperature_Min = mean(`Temperature Min`, na.rm = TRUE),
    Avg_Temperature_Max = mean(`Temperature Max`, na.rm = TRUE),
    Avg_pH_Min = mean(`pH Min`, na.rm = TRUE),
    Avg_pH_Max = mean(`pH Max`, na.rm = TRUE),
    Avg_Conductivity_Min = mean(`Conductivity (µmhos/cm) Min`, na.rm = TRUE),
    Avg_Conductivity_Max = mean(`Conductivity (µmhos/cm) Max`, na.rm = TRUE),
    .groups = 'drop' 
  )

print(yearly_data)
```


```{r}
station_freq=table(df$`Station Name`)
state_freq=table(df$STATE)

station_freq_df = as.data.frame(station_freq)
state_freq_df = as.data.frame(state_freq)

colnames(station_freq_df) = c("Station Name", "Frequency")
colnames(state_freq_df) = c("State", "Frequency")

print(station_freq_df)
print(state_freq_df)
```


```{r}
hist(df$`Temperature Min`, main = "Histogram of Temperature Min", xlab = "Temperature Min", breaks = 20, col = "blue")
hist(df$`Temperature Max`, main = "Histogram of Temperature Max", xlab = "Temperature Max", breaks = 20, col = "green")
hist(df$`pH Min`, main = "Histogram of pH Min", xlab = "pH Min", breaks = 20, col = "red")
hist(df$`pH Max`,main = "Histogram of pH Max",xlab = "pH Max", breaks = 20, col = 'orange')
hist(df$`Conductivity (µmhos/cm) Min`, main = "Histogram of Conductivity Min", xlab = "Conductivity Min", breaks = 20, col = "purple")
hist(df$`Conductivity (µmhos/cm) Max`,main = "Histogram of Conductivity Max", xlab = "Conductivity Max", breaks = 20, col = 'yellow')




```

The histogram of minimum temperature appears to be right-skewed
The histogram of maximum temperature appears to be right-skewed
The histogram of pH Min appears to be approximately normal
The histogram of pH Max appears to be right-skewed
The histogram of Conductivity Min appears to be right-skewed
The histogram of Conductivity Max appears to be right-skewed. 

```{r}
library(dplyr)

outlier_detection <- df %>%
  group_by(STATE) %>%
  mutate(
    Temp_Min_IQR = IQR(`Temperature Min`, na.rm = TRUE),
    Q1_Temp_Min = quantile(`Temperature Min`, 0.25, na.rm = TRUE),
    Q3_Temp_Min = quantile(`Temperature Min`, 0.75, na.rm = TRUE),

    Temp_Max_IQR = IQR(`Temperature Max`, na.rm = TRUE),
    Q1_Temp_Max = quantile(`Temperature Max`, 0.25, na.rm = TRUE),
    Q3_Temp_Max = quantile(`Temperature Max`, 0.75, na.rm = TRUE),
    
    pH_Max_IQR = IQR(`pH Max`, na.rm = TRUE),
    Q1_pH_Max = quantile(`pH Max`, 0.25, na.rm = TRUE),
    Q3_pH_Max = quantile(`pH Max`, 0.75, na.rm = TRUE),
    
    Conductivity_Min_IQR = IQR(`Conductivity (µmhos/cm) Min`, na.rm = TRUE),
    Q1_Conductivity_Min = quantile(`Conductivity (µmhos/cm) Min`, 0.25, na.rm = TRUE),
    Q3_Conductivity_Min = quantile(`Conductivity (µmhos/cm) Min`, 0.75, na.rm = TRUE),
    
    Conductivity_Max_IQR = IQR(`Conductivity (µmhos/cm) Max`, na.rm = TRUE),
    Q1_Conductivity_Max = quantile(`Conductivity (µmhos/cm) Max`, 0.25, na.rm = TRUE),
    Q3_Conductivity_Max = quantile(`Conductivity (µmhos/cm) Max`, 0.75, na.rm = TRUE),

    
    pH_Min_Z = (`pH Min` - mean(`pH Min`, na.rm = TRUE)) / sd(`pH Min`, na.rm = TRUE)
  ) %>%
 
  mutate(
    Temp_Min_Outlier = `Temperature Min` < (Q1_Temp_Min - 1.5 * Temp_Min_IQR) | `Temperature Min` > (Q3_Temp_Min + 1.5 * Temp_Min_IQR),
    Temp_Max_Outlier = `Temperature Max` < (Q1_Temp_Max - 1.5 * Temp_Max_IQR) | `Temperature Max` > (Q3_Temp_Max + 1.5 * Temp_Max_IQR),
    pH_Max_Outlier = `pH Max` < (Q1_pH_Max - 1.5 * pH_Max_IQR) | `pH Max` > (Q3_pH_Max + 1.5 * pH_Max_IQR),
    Conductivity_Min_Outlier = `Conductivity (µmhos/cm) Min` < (Q1_Conductivity_Min - 1.5 * Conductivity_Min_IQR) | `Conductivity (µmhos/cm) Min` > (Q3_Conductivity_Min + 1.5 * Conductivity_Min_IQR),
    Conductivity_Max_Outlier = `Conductivity (µmhos/cm) Max` < (Q1_Conductivity_Max - 1.5 * Conductivity_Max_IQR) | `Conductivity (µmhos/cm) Max` > (Q3_Conductivity_Max + 1.5 * Conductivity_Max_IQR),

    pH_Min_Outlier = abs(pH_Min_Z) > 3
  )

outlier_detection

```

```{r}
colnames(outlier_detection)
```

```{r}
outliers <- outlier_detection %>%
  filter(Temp_Min_Outlier == TRUE | 
         Temp_Max_Outlier == TRUE | 
         pH_Max_Outlier == TRUE | 
         Conductivity_Min_Outlier == TRUE | 
         Conductivity_Max_Outlier == TRUE | 
         pH_Min_Outlier == TRUE)

head(outliers)

```

```{r}
outlier_count_by_state <- outliers %>%
  group_by(STATE) %>%
  summarise(Outlier_Count = n())

print(outlier_count_by_state)

```

```{r}

yearly_data <- df %>%
  group_by(Year) %>%
  summarise(Avg_Temperature_Min = mean(`Temperature Min`, na.rm = TRUE))

temperature_ts <- ts(yearly_data$Avg_Temperature_Min, 
                     start = min(yearly_data$Year), 
                     end = max(yearly_data$Year), 
                     frequency = 1)

# Apply first differencing to the time series
diff_temperature_ts <- diff(temperature_ts)

# Plot the differenced series
plot(temperature_ts, main = "Temperature Time Series", ylab = "Avg Temperature Min", xlab = "Year")


```
```{r}
library(tseries)
adf_test_diff <- adf.test(diff_temperature_ts)
print(adf_test_diff)


```
```{r}
library(forecast)
# Fit ARIMA model on the original series with first differencing
arima_model_diff <- auto.arima(temperature_ts, d = 1)  # d=1 for first differencing

# Summary of the ARIMA model
summary(arima_model_diff)

```
```{r}
# Forecast the next 5 periods (e.g., years)
forecast_diff_values <- forecast(arima_model_diff, h = 20)

# Plot the forecast
plot(forecast_diff_values, main = "ARIMA Forecast for Temperature Min", xlab = "Year", ylab = "Temperature Min")

# Print the forecasted values
print(forecast_diff_values)

```
```{r}
# Check residuals of the ARIMA model
checkresiduals(arima_model_diff)

```

```{r}
# Assuming df contains data up to 2021
library(dplyr)

# Prepare the training data
train_data <- df %>% filter(Year <= 2021)

# Create a linear model for each target variable
lm_models <- list()
target_columns <- c("Temperature Min", "Temperature Max", "pH Min", "pH Max", 
                    "Conductivity (µmhos/cm) Min", "Conductivity (µmhos/cm) Max")

# Train models for each target variable
for (target in target_columns) {
  formula <- as.formula(paste("`", target, "` ~ Year", sep = ""))  # Enclose column name in backticks
  lm_models[[target]] <- lm(formula, data = train_data)
}

# Forecast for 2022–2026
forecast_years <- 2022:2030
forecasted_values <- data.frame(Year = forecast_years)

# Apply the models to forecast future values
for (target in target_columns) {
  forecasted_values[[target]] <- predict(lm_models[[target]], newdata = forecasted_values)
}

# Print the forecasted values for 2022–2026
print(forecasted_values)
```

```{r}
# Assuming df contains data up to 2021
library(dplyr)
library(ggplot2)
library(tidyr)

# Prepare the training data
train_data <- df %>% filter(Year <= 2021)

# Create a linear model for each target variable
lm_models <- list()
target_columns <- c("Temperature Min", "Temperature Max", "pH Min", "pH Max", 
                    "Conductivity (µmhos/cm) Min", "Conductivity (µmhos/cm) Max")

# Train models for each target variable
for (target in target_columns) {
  formula <- as.formula(paste("`", target, "` ~ Year", sep = ""))  # Enclose column name in backticks
  lm_models[[target]] <- lm(formula, data = train_data)
}

# Forecast for 2022–2026
forecast_years <- 2022:2030
forecasted_values <- data.frame(Year = forecast_years)

# Apply the models to forecast future values
for (target in target_columns) {
  forecasted_values[[target]] <- predict(lm_models[[target]], newdata = forecasted_values)
}

# Reshape the forecasted data for plotting (long format)
forecasted_values_long <- forecasted_values %>%
  gather(key = "Variable", value = "Forecast", -Year)  # Using gather() to reshape

# Plot the forecasted values using ggplot2 and facet_wrap
ggplot(forecasted_values_long, aes(x = Year, y = Forecast, color = Variable, group = Variable)) +
  geom_line(size = 1) +
  geom_point(size = 3) +
  labs(title = "Forecasted Values for 2022–2030",
       x = "Year",
       y = "Forecasted Value") +
  theme_minimal() +
  theme(legend.position = "bottom") +
  scale_color_manual(values = c("Temperature Min" = "blue", 
                                "Temperature Max" = "red", 
                                "pH Min" = "green", 
                                "pH Max" = "purple", 
                                "Conductivity (µmhos/cm) Min" = "orange", 
                                "Conductivity (µmhos/cm) Max" = "brown")) +
  facet_wrap(~ Variable, scales = "free_y")  # Creates separate plots for each variable


```








